Automatically generated NE tagged corpora for English and Hungarian

### 2012-03-25 paper/engNERwiki.tex
ezmegaz

The types in DBpedia are organized into a class hierarchy, available as an OWL\footnote{http://www.w3.org/TR/owl-ref/} ontology. The ontology contains the 320 most frequent entity categories, arranged into a taxonomy under the base class \texttt{owl:Thing}. Most of the classes belong into the 6 largest sub-hierarchies (here named after their topmost classes): \texttt{Person}, \texttt{Organisation}), \texttt{Event}, \texttt{Place}, \texttt{Species} and \texttt{Work}. The taxonomy is rather flat: the top level contains 44 classes and there are several nodes with a branching factor of 20.

@@

From the available NER annotation standards we elected to use the CoNLL \cite{Tjong:03} NE types. It is not difficult to see the parallels between the DBpedia sub-hierarchies \texttt{Person}, \texttt{Organisation} and \texttt{Place} and the CoNLL NE categories PERSON, ORGANIZATION and LOCATION. The fourth category, MISCELLANEOUS is more elusive; according to the CoNLL NER annotation guide\footnote{http://www.cnts.ua.ac.be/conll2003/ner/annotation.txt}, the sub-hierarchies \texttt{Event} and \texttt{Work} belong to this category, as well as various other classes outside the main hierarchies.

@@

Using the mapping described above, we compile a NE list that contains all entities in DBpedia tagged with the appropriate CoNLL category.

@@

A NER corpus consists of sentences where NEs are accurately tagged. In order to automatically prepare such sentences, two tasks need to be performed: identifying the entities in the sentence and tagging them with the correct tag. Sentences, for which accurate tagging could not be accomplished, must be removed from the corpus.

Our approach to entity identification relies on Wikipedia cross-references found in the article text. We assume that individual Wikipedia articles describe NEs. A link to an article can then be perceived as a mapping that identifies its anchor text with a particular entity.

Discovered entities are tagged with the CoNLL category assigned to them in the entity list extracted from DBpedia. If the link target is not in the entity list, or the link points to a disambiguation page, we cannot determine the type of the entity and tag it as \texttt{UNK}. Links to redirect pages are resolved to point instead to the redirect target, after which they are handled as regular cross-references. Finally, sentences with UNK links in them are removed from the corpus.

The following sub-sections describe how the method explained above can be improved to increase precision, sentence coverage and to account for peculiarities in the CoNLL guidelines.

@@

We evaluated our English corpus against (a) itself (divided into train and test sets); and (b) a manually annotated English corpus. Since the filtered English Wikipedia corpus -- containing only the sentences with NEs -- is still very huge, and we are limited by time and memory, our experiments were performed with a portion of 3.5 million tokens (the size of our filtered Hungarian corpus). It was divided into train and test sets (90\%-10\%).

For the English cross-corpus evaluation the CoNLL-2003 dataset was chosen, mostly because of the similarity in tagset and format. It is well-studied that cross-checking between different corpora leads to F-measure decrease. Stylistic and domain differences certainly affect NER performance, and the different annotation schemes pose several compatibility problems. Nothman et al.~\shortcite{Nothman:08} showed that each set of gold standard training data performs much better on corresponding test sets than on test sets from other sources. The situation here is similar: the NE-tagger trained on Wikipedia does not achieve as high performance tested against CoNLL test set as one trained on the own training set (see Table \ref{enresults} for results).

>>>>

The types in DBpedia are organized into a class hierarchy, available as an OWL\footnote{http://www.w3.org/TR/owl-ref/} ontology. The ontology contains the 320 most frequent entity categories, arranged into a taxonomy under the base class \texttt{owl:Thing}. Most of the classes belong into the 6 largest sub-hierarchies (here named after their topmost classes): \texttt{Person}, \texttt{Organisation}, \texttt{Event}, \texttt{Place}, \texttt{Species} and \texttt{Work}. The taxonomy is rather flat: the top level contains 44 classes and there are several nodes with a branching factor of 20.

@@

From the available NER annotation standards we elected to use the CoNLL \cite{Tjong:03} NE types. It is not difficult to see the parallels between the DBpedia sub-hierarchies \texttt{Person}, \texttt{Organisation} and \texttt{Place} and the CoNLL NE categories PER, ORG and LOC. The fourth category, MISC is more elusive; according to the CoNLL NER annotation guide\footnote{http://www.cnts.ua.ac.be/conll2003/ner/annotation.txt}, the sub-hierarchies \texttt{Event} and \texttt{Work} belong to this category, as well as various other classes outside the main hierarchies.

@@

Using the mapping described above, we compile a list that contains all entities in DBpedia tagged with the appropriate CoNLL category.

@@

A NER corpus consists of sentences where NEs are accurately tagged. In order to automatically prepare such sentences, two tasks need to be performed: identifying entities in the sentence and tagging them with the correct tag. Sentences, for which accurate tagging could not be accomplished, must be removed from the corpus.

Our approach to entity identification relies on Wikipedia cross-references found in the article text. We assume that individual Wikipedia articles describe NEs. A link to an article can then be perceived as a mapping that identifies its anchor text with a particular NE.

Discovered entities are tagged with the CoNLL label assigned to them in the entity list extracted from DBpedia. If the link target is not in the entity list, or the link points to a disambiguation page, we can not determine the type of the entity, and tag it as UNK. Links to redirect pages are resolved to point instead to the redirect target, after which they are handled as regular cross-references. Finally, sentences with UNK links in them are removed from the corpus.

There are several ways to increase precision, sentence coverage and to account for peculiarities in the CoNLL guidelines. These are detailed in the next subsections.

@@

We evaluated our English corpus against (a) itself (divided into train and test sets); and (b) a manually annotated English corpus. Since the filtered English Wikipedia corpus, containing only the sentences with NEs, is still very huge, and we are limited by time and memory, our experiments were performed with a portion of 3.5 million tokens (the size of our filtered Hungarian corpus). It was divided into train and test sets (90\%-10\%).

For the English cross-corpus evaluation the CoNLL-2003 dataset was chosen, mostly because of the similarity in tagset and format. It is well-studied that cross-checking between different corpora decrease F-measure. Stylistic and domain differences certainly affect NER performance, and the different annotation schemes pose several compatibility problems. Nothman et al.~\shortcite{Nothman:08} showed that each set of gold standard training data performs much better on corresponding test sets than on test sets from other sources. The situation here is similar: the NE-tagger trained on Wikipedia does not achieve as high performance tested against CoNLL test set as one trained on the own training set (see Table \ref{enresults} for results).
